A Switching Planner for Combined Task and Observation Planning

### 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
abstract
More DTPDDL

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
(a.k.a., relational) language that facilitates factored description of
domains and problem instances. There are straightforward compilations
from problems expressed compactly in PPDDL to propositional
representations amenable to state-of-the-art planning procedures.
Because this language cannot model domains that feature partial
observability, for our setting we have developed DTPDDL, an extension
of PPDDL to partial observability.

Our mobile robot is able to look for a visual object, for example a
box of cornflakes, at a given place. We can model that deterministic
action using the following schema:

\def\oom{$^{\tt +}$}
\def\zom{$^*$}
\def\bump{\hspace{1cm}}
\def\req#1{$^{\tt #1}$}
\def\noteme#1{}%{[[#1]]}
\def\notecoauth#1{\ }
\def\notereader#1{[[#1]]}
\def\meta{$\uparrow\uparrow$}
\def\la{\langle}
\def\ra{\rangle}

\small{
\begin{tabtt}
(\= :action look-for-object \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - location) \\
  \> :precondition (and (= (is-in ?r) ?l) ) \\
  \> :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
}

\noindent That is, the robot can always look for a visual object at
its current location. The only effect this has on the state, is that
the robot incurs a penalty, $3$, that corresponds to the {\em cost} of
performing a visual search. Instances of this operator are useful,
because their application activates an the following sense schema:

\small{
\begin{tabtt}
(\= :sense vision \\
  \> :parameters \=(?r - robot ?v - visual-object\\
  \> \> ?l - location) \\
  \> :execution (look-for-object ?r ?v ?l) \\
  \> :effect ( and (when (= (is-in ?v) ?l) \\
  \> \bump (probabilistic 0.8 \\
  \> \bump (= (observe-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
  \> \bump (probabilistic 0.1 \\
  \> \bump (= (observe-is-in ?v) ?l))))) \\
\end{tabtt}
}

\noindent Here, the {\em execution} clause is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. No
other conditions guard the activation of this sense, therefore if
$(\pp{look-for-object} \pp{Dora} \pp{cornflakes} \pp{bedroom})$ is
executed, we suppose there is a $0.8$ chance of observing the
cornflakes if they are in the $\pp{bedroom}$, and otherwise a $0.1$ of
observing them.\footnote{We have simplified the schemata for
illustrative purposes. In practice, the probabilities of a particular
observation here should be parametrised by the type of visual object
the robot is searching for, and category of the location.}

\scriptsize
\begin{nopagebreak}\begin{tabtt}
<observation-def> \=::= \=(:sense <observation symbol> \\
                        \> \> :parameters (<typed list (variable)>)  \\
                        \> \> <o-def body>) \\
  <o-symbol> \> ::= <name> \\
  <o-def body> \> ::= [:precondition <GD>] \\
  \> \> [:execution <atomic action(term)> ] \\
    \> \> [:effect <o-effect>] \\
  <atomic action(t)> \> ::= (<action symbol> t\zom) \\
  <o-effect> \> ::= (and <c-o-effect>\zom) \\
  <o-effect> \> ::= <c-o-effect> \\
  <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
  <c-o-effect> \> ::= <p-o-effect> \\
  <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
  <p-o-effect> \> ::= <atomic o-formula(term)> \\
  <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
  <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
  <o-f-exp> \> ::= <number>\\
  <o-f-exp> \> ::= (- <o-f-exp>)\\
  <o-f-exp> \> ::= <o-f-head>\\
  <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
  <o-f-head> \> ::= <o-function-symbol>\\
\end{tabtt}\end{nopagebreak}
\noteme{<structure-def> ::= <attach-def>}
\normalsize

@@

For example, $(:\pp{predicates} (\pp{location} ?l - VisualObject))$
says that there is a unary predicate

knowledge objectives and  sensory effects.

probabilistic models of the sensing consequences of acting to
quantitatively capture the unreliability of sensing, and thus reason
precisely about the utility of a plan.

@@

%%
Recently the system
\system{Conformant-FF}~\cite{hoffmann:brafman:2006}, demonstrated how
conformant planning ---i.e., serial planning in unobservable worlds---
can be modelled as a deterministic problem, and therefore solved using
serial systems. In this conformant setting, advances have been towards
compact representations and lazy evaluations of belief-states.

@@

>>>>

Realistic robot planning problems in uncertain environments often
require achieving tasks while also finding out about the
world. Because the world state cannot be observed directly, these
problems are naturally represented as partially observable Markov
decision problems (POMDPs). However, these are typically intractable
for realistic problems.

We present a \emph{switching planner} that employs fast sequential
planning to decide on the overall strategy, and uses a
decision-theoretic planner to solve the subproblems where partial
observability will significantly impact the quality of the plan. We
demonstrate the approach in a realistic robot exploration domain.

@@

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
(a.k.a., relational) language that facilitates factored description of
domains and problem instances. There are straightforward compilations
from problems expressed compactly in PPDDL to propositional
representations amenable to state-of-the-art planning procedures.

Because PPDDL cannot model domains that feature partial observability,
we developed an extension of that language, called DTPDDL. This can
express probabilistic models of the sensing consequences of acting to
quantitatively capture unreliability in perception. That expressive
power is achieved by incorporating perceptual analogues of fluent and
predicate signatures, and of action definitions. In detail, we have
declarations of state characterising predicate and fluent symbols as
usual. In addition, we allow two other declarative blocks, for
perceptual predicates and fluents respectively. In detail, suppose
our robot is tasked with exploring {\em locations} in order to
identify the whereabouts of a {\em visual-object}. We must describe
state and perceptual facts that model the {\em true}, resp. perceived,
locations of objects. In DTPDDL, these declarations appear as:

\small{
\begin{tabtt}
(:\=functions  ;; state fluents\\
  \> (is-in ?v - visual-object) - location )\\
(:observable-functions  ;; perceptual fluents\\
  \> (o-is-in ?v - visual-object) - location )
\end{tabtt}
}

\noindent To model the sensing capabilities of the agent, we have
operator-like descriptive elements, with preconditions in both action
and state symbols, and uniformly positive effects over perceptual
predicates -- I.e, the negation normal form of a sensing effect
formula cannot contain negation. Moreover, and without a loss of
generality, we find it convenient to separate the actional
precondition from the state precondition. For example, our mobile
robot is able to look for a visual object, such as a box of
cornflakes, at a given place. We can model that deterministic action
using the following operator schema:

\small{
\begin{tabtt}
(\=:action look-for-object \+ \\
   :parameters (\=?r - robot ?v - visual-object\\
   \> ?l - location) \\
   :precondition (and (= (is-in ?r) ?l) ) \\
   :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
}

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect effect on the state. That is,
execution incurs a penalty, $3$, corresponding to the {\em cost} of
performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
reserved word}, and occurs in {\em increased} and {\em decreased}
clauses of operator schemata. In that setting, {\em reward} is
interpreted as accumulated rewards and penalties.}

Completing the example, we model the sensing outcome that results from
instantiating the $\pp{look-for-object}$ operator:

\small{
\begin{tabtt}
(\= :sense vision \+\\
 :parameters \= (\= ?r - robot ?v - visual-object\\
 \>\>  ?l - location) \\
 :execution \> ( \> look-for-object ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
   \> \> (probabilistic 0.8 \\
   \>  \>(= (observe-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
   \>  \> (probabilistic 0.1 \\
   \>  \> (= (observe-is-in ?v) ?l))))) \\
\end{tabtt}
}

\noindent Here, the {\em execution} clause is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. We
also include a redundant state precondition, $(=(\pp{is-in}~?r)?l)$,
in order to fully demonstrate the syntax. Interpreting the schema, if
action
$(\pp{look-for-object}~\pp{robot}~\pp{cornflakes}~\pp{bedroom})$ is
executed, there is a $0.8$ chance of perceiving the cornflakes if they
are in the $\pp{bedroom}$, and otherwise a $0.1$ of perceiving them.

%% \footnote{We have simplified the schemata for illustrative
%% purposes. In practice, the probabilities of a particular observation
%% here should be parametrised by the type of visual object the robot is
%% searching for, and category of the location.}

@@

@@

%%
Recently the systems
\system{Conformant-FF}~\cite{hoffmann:brafman:2006} and
$T_0$~\cite{palacios:geffner:2009}, demonstrated how conformant
planning ---i.e., serial planning in unobservable worlds--- can be
modelled as a deterministic problem, and therefore solved using serial
systems. In this conformant setting, advances have been towards
compact representations of beliefs amenable to existing best-first
search planning procedures, and lazy evaluations of beliefs. This
research thread has also been extended recently to contingent planning
in fully observable non-deterministic
environments~\cite{albore:etal:2009}.

@@

%% \scriptsize
%% \begin{nopagebreak}\begin{tabtt}
%% <observation-def> \=::= \=(:sense <observation symbol> \\
%%                         \> \> :parameters (<typed list (variable)>)  \\
%%                         \> \> <o-def body>) \\
%%   <o-symbol> \> ::= <name> \\
%%   <o-def body> \> ::= [:precondition <GD>] \\
%%   \> \> [:execution <atomic action(term)> ] \\
%%     \> \> [:effect <o-effect>] \\
%%   <atomic action(t)> \> ::= (<action symbol> t\zom) \\
%%   <o-effect> \> ::= (and <c-o-effect>\zom) \\
%%   <o-effect> \> ::= <c-o-effect> \\
%%   <c-o-effect> \> ::= \req{:probabilistic-effects} (probabilistic <prob> <o-effect>) \\
%%   <c-o-effect> \> ::= <p-o-effect> \\
%%   <atomic o-formula(t)> \> ::= (<observation> t\zom) \\
%%   <p-o-effect> \> ::= <atomic o-formula(term)> \\
%%   <p-o-effect> \> ::= (not <atomic o-formula(term)>) \\
%%   <o-f-comp> \> ::= (<binary-comp> <o-f-exp> <o-f-exp>)\\
%%   <o-f-exp> \> ::= <number>\\
%%   <o-f-exp> \> ::= (- <o-f-exp>)\\
%%   <o-f-exp> \> ::= <o-f-head>\\
%%   <o-f-head> \> ::= (<o-function-symbol> <term>\zom )\\
%%   <o-f-head> \> ::= <o-function-symbol>\\
%% \end{tabtt}\end{nopagebreak}
%% \noteme{<structure-def> ::= <attach-def>}
%% \normalsize
